---
title: "An exploration of Halloween candy sales."
output:
  html_document:
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
---

###Purpose
To examine candy sales patterns in Canada and the USA and explore a limited set of potential consequences.

### Load Libraries
```{r}
library(tidyverse)
```

### Load Data
```{r}
canCandy <-read.csv("data/canadian.candy.sales.csv")
humanPop <-read.csv("data/human.pop.density.csv")
usaCandy <- read.csv("data//usa.halloween.spending.csv")
```

### Questions
1. How have Halloween candy sales changed over time?
2. How much does Halloween candy sales relate to yearly candy sales?
3. Is October the best month for candy sales?
4. Are current candy sales "profitable"?


### EDA
```{r}
## Using Tibble
as_data_frame(canCandy)

## Using compact display
str(canCandy)

## others views
head(canCandy)

## see change in sales over years
ggplot(canCandy) +geom_boxplot(mapping = aes(year, candy.sales)) + coord_flip()

## need as a factor
canCandy$year <-as.factor(canCandy$year)
str(canCandy)

## plot again with factor
ggplot(canCandy) +geom_boxplot(mapping = aes(year, candy.sales)) + coord_flip()

## see the difference in sales over months
canCandy$month <-as.factor(canCandy$month)
ggplot(canCandy) +geom_boxplot(mapping = aes(month, candy.sales)) + coord_flip()


## see the difference in sales in just October
octCandy <- canCandy %>% filter(month == 10) 

## plot change in October
ggplot(octCandy) +geom_point(mapping = aes(year, candy.sales))

## Clean up plot
ggplot(octCandy) +geom_point(mapping = aes(year, candy.sales)) + ylab("Candy Sales ($)") + xlab("Year") + theme_classic()
```



### Fitting a linear model with October Candy Sales
```{r}
## check distribution of data
hist(octCandy$candy.sales)



## fit a linear model
m1 <- lm(candy.sales ~ as.numeric(year), data= octCandy)
summary(m1) 

## check model fit
hist(m1$residuals)
shapiro.test(m1$residuals)

## fit a linear model
m2 <- lm(log(candy.sales) ~ as.numeric(year), data= octCandy)
summary(m2) 

## check model fit
hist(m2$residuals)
shapiro.test(m2$residuals)

## Plot again but log-transformed
ggplot(octCandy) +geom_point(mapping = aes(year, log(candy.sales))) + ylab("Candy Sales ($)") + xlab("Year") + theme_classic()

## Plot again but corrected y-axis
ggplot(octCandy, aes(as.numeric(year), candy.sales)) +geom_point() + ylab("Candy Sales ($)") + xlab("Year")  +  scale_y_log10(limits = c(100000,500000)) + geom_smooth(method=lm)

anova(m2, test="Chisq")

```

### Fitting a linear model with average yearly Candy
```{r}
avgCandy <- canCandy %>% group_by(year) %>%  summarize(sales=mean(candy.sales),error=sd(candy.sales))

hist(avgCandy$sales)

## fit a linear model 
m3 <- lm(sales ~ as.numeric(year), data=avgCandy)
summary(m3)

## check normality
shapiro.test(m3$residuals)


## fit a linear model with log transformed
m4 <- lm(log(sales) ~ as.numeric(year), data=avgCandy)
summary(m4)

## check normality
shapiro.test(m4$residuals)

```

### Compare October sales to annual sales
```{r}
## Calculate average yearly sales
annual.sales <-canCandy %>% group_by(year) %>% summarise(totalYear = sum(candy.sales))
annual.sales

## divide october sales by year
octCandy <- octCandy %>% mutate(weighted = candy.sales/annual.sales$totalYear)
octCandy

## make it a percentage
octCandy <- octCandy %>% mutate(weighted = candy.sales/annual.sales$totalYear*100)
octCandy

## plot
ggplot(octCandy, aes(year, weighted)) + geom_point()

## check if there is a trend over time
m1 <- lm(weighted ~ as.numeric(year), data=octCandy)
summary(m1)

## test normality
shapiro.test(m1$residuals)

## plot
ggplot(octCandy, aes(as.numeric(year), weighted)) + geom_point() + geom_smooth(method="lm") + ylab("Proportion of October sales for the year (%)")

## A polynomial?
ggplot(octCandy, aes(as.numeric(year), weighted)) + geom_point() + geom_smooth(method="gam", formula= y ~ s(x))+ ylab("Proportion of October sales for the year (%)")

```


### check the best month over years
```{r}

totalMonth <- canCandy %>% group_by(year, month) %>% summarize(sales=sum(candy.sales))

## best month
totalMonth %>% group_by(year) %>% slice(which.max(sales))

## worst month
totalMonth %>% group_by(year) %>% slice(which.min(sales))
```

### Are candy sales profitable?
```{r}
## create column for cost to stay in business
totalMonth <- totalMonth %>%  mutate(profit = ifelse(sales > 200000, 1, 0))

## check the profit over time
ggplot(totalMonth) + geom_bar(aes(x=year, y=profit), stat="identity")


## GLM to determine if the change is significant
m1 <- glm(profit ~ as.numeric(year), family="binomial", data=totalMonth)
summary(m1)

anova(m1, test="Chisq")

```





